---
title: "Comparison of Pollen Traps"
subtitle: "Evaluation of Similarity and Robustness of Eight Pollen Traps Located in Payerne During the Blooming Season 2019"
author: "Simon Adamov"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
always_allow_html: TRUE
output:
  html_document:
    df_print: paged 
  pdf_document: default
  word_document: default
---

# Setup

This project is using renv dependency management, for more info: https://cran.r-project.org/web/packages/renv/vignettes/renv.html
The .RProfile is optimized for an interactive session in VSCode.

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  fig.width = 10,
  fig.height = 7,
  out.width = "100%",
  out.height = "100%"
)
# This project is using renv dependency management, for more info:
# https://cran.r-project.org/web/packages/renv/vignettes/renv.html

library(caTools)
library(MASS)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(here)
library(kableExtra)
library(ggthemr)

# library(AeRobiology)
# library(lme4)
# library(psych)
# library(robustlmm)
# library(nparcomp)
# library(goftest)

library(conflicted)
# conflict_prefer()

# Some functions are used from these packages
# missmap from library(Amelia)
# transpose from library(data.table)
# statistical test from library(dunn.test)
# statistical test from library(conover.test)

devtools::load_all()
# I like the look of these plots: cttobin/ggthemr
ggthemr("fresh")

# Due to old R-Version some packages must be installed from CRAN Archive,
# in order to knit markdown documents.
# caTools@1.17.1.1
# pbkrtest@0.4-7
# nloptr@1.2.2
# foreign@0.8-76
# devtools@2.2.1
# usethis@1.5.1
# devtools@2.2.1
```

# Data Import

The data was prepared by Fiona Tummon and stored in the /data-raw Folder.
Eight different Pollen Traps were situated on the roof in Payerne and continously measured pollen grains between 18.04.2019 - 04.06.2019.
For the sake of simplicity we only look at concentrations of total pollen (i.e. the sum of all pollen taxa).
The concentrations were averages to obtain three different timeseries:

- Hourly averages
- 6-hour averages (hourly values from 00:00 to 05:00, 06:00 to 11:00, ...)
- Daily averages (hourly values from 00:00 to 23:00)

The preprocessed data is available to the reader in the /data folder. In /data-raw/preproc.R the preprocessing is documented.
The data has no missing or erroneous data, as it was carefully prepared by Fiona.

```{r }

```